### 2017-08-16 paper/paper.Rmd
update JOSS paper and adjust figures

---
title: 'visdat: Visualising Whole Data Frames'
authors:
- affiliation: 1
  name: Nicholas Tierney
  orcid: 0000-0003-1460-8722
date: "01 August 2017"
output:
  pdf_document: default
  html_document:
    keep_md: yes
bibliography: paper.bib
tags:
- visualisation
- R
- exporatory data analysis
affiliations:
- index: 1
  name: Monash University
---

JOSS requirements:

- A summary describing the high-level functionality and purpose of the software for a diverse, non-specialist audience
- A clear statement of need that illustrates the purpose of the software
- A list of key references including a link to the software archive
- Mentions (if applicable) of any ongoing research projects using the software or recent scholarly publications enabled by it

# Summary

When you receive a new dataset you need to look at the data to get a sense of what is in it and understand potential problems and challenges to get it analysis-ready. "Taking a look at the data" can mean different things. For example, examining statistical summaries (minimum, maximum, mean, interquartile range), finding missing values, checking data formatting, creating graphical summaries such as histograms, scatterplots, boxplots, and more.

When handling typical "messy" data, these preliminary exploratory steps can become difficult to perform when you encounter data that is not what you expect. For example, income might be a factor instead of numeric, date could be a number not a character string or a date class, or values could be missing when they shouldn't be. Often times you realise that you have expectations of the data, but it can be hard to know what they are exactly until they are a problem. This is sort of like how you might not think to buy more lightbulbs until one goes out - when you go to use the data in an exploratory scatter plot, or in a preliminary model, you don't always realise your data is in the wrong format until that moment. What you need is a birds eye view of the data that tells you what classes are in the dataframe, and where the missing data are.

`visdat` is an R [@Rcore] package that provides a tool to "get a look at the data", creating heatmap-like visualisations of an entire dataframe - classes in the data, missing values, and comparisons between two datasets. `visdat` takes inspiration from [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), and is powered by ggplot2 [@ggplot2], which provides a consistent and powerful framework for visualisations that can be extended if needed.

Plots are presented in an intuitive way, reading top down, just like your data. Below is a plot of some example typical data containing missing values and data of a variety of classes.

```{r load-data}
library(visdat)
vis_dat(typical_data)

```

`visdat` will continue to be improved over time, to improve speed in computation and also add interactive plots.

# Acknowlegements

I would like to thank the two reviewers, Mara Averick and Sean Hughes, for their helpful suggestions that resulted in a much better package, and rOpenSci for providing the support of the onboarding package review that facilitated these improvements.

# References