### 2018-06-06 README.Rmd
use project status badge instead of lifecycle

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

```

# visdat <img src="man/figures/visdat-logo.png" align="right" />

[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/onboarding/issues/87)[![JOSS status](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a/status.svg)](http://joss.theoj.org/papers/c85f57adbc565b064fb4bfc9b59a1b2a)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)

[![Travis Build Status](https://travis-ci.org/ropensci/visdat.svg?branch=master)](https://travis-ci.org/ropensci/visdat)[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/ropensci/visdat?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/visdat)[![Coverage Status](https://img.shields.io/codecov/c/github/ropensci/visdat/master.svg)](https://codecov.io/github/ropensci/visdat?branch=master)

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/naniar)](https://cran.r-project.org/package=naniar)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](http://cran.rstudio.com/web/packages/visdat/index.html)

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

# How to install

visdat is available on CRAN

```{r install-cran, eval = FALSE}

install.packages("visdat")

```

If you would like to use the development version, install from github with:

```{r installation, eval = FALSE}

# install.packages("devtools")
devtools::install_github("ropensci/visdat")
```

# What does visdat do?

Initially inspired by [`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `vis_dat` helps you visualise a dataframe and "get a look at the data" by displaying the variable classes in a dataframe as a plot with `vis_dat`, and getting a brief look into missing data patterns using `vis_miss`.

The name `visdat` was chosen as I think in the future it could be integrated with [`testdat`](https://github.com/ropensci/testdat). The idea being that first you visualise your data (`visdat`), then you run tests from `testdat` to fix them.

There are two main commands in the `visdat` package:

- `vis_dat()` visualises a dataframe showing you what the classes of the columns are, and also displaying the missing data.

- `vis_miss()` visualises just the missing data, and allows for missingness to be clustered and columns rearranged. `vis_miss()` is similar to `missing.pattern.plot` from the [`mi`](https://cran.r-project.org/web/packages/mi/index.html) package. Unfortunately `missing.pattern.plot` is no longer in the `mi` package (as of 14/02/2016).

You can read more about visdat in the vignette, ["using visdat"](http://visdat.njtierney.com/articles/using_visdat.html).

There are also experimental functions, which include guessing values,  comparing two dataframes, and adding interactivity. These are described in detail in the vignette ["Experimental Features of visdat"](http://visdat.njtierney.com/articles/experimental_features.html).

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

# Examples

## Using `vis_dat()`

Let's see what's inside the `airquality` dataset from base R, which contains information about daily air quality measurements in New York from May to September 1973. More information about the dataset can be found with `?airquality`.

```{r vis-dat-aq}

library(visdat)

vis_dat(airquality)

```

The plot above tells us that R reads this dataset as having numeric and integer values, with some missing data in `Ozone` and `Solar.R`. The classes are represented on the legend, and missing data represented by grey. The column/variable names are listed on the x axis.

By default, `vis_dat` sorts the columns according to the type of the data in the vectors. You can turn this off by setting `sort_type = FALSE`.

```{r vis-dat-aq-sort-type}

vis_dat(airquality, 
        sort_type = FALSE)

```

### With many kinds of data

To demonstrate what visdat looks like when you have different kinds of data, we can look at the dataset `typical_data`, provided within `visdat`, and created with the excellent [`wakefield` package](https://github.com/trinker/wakefield).

```{r vis-dat-typical-data}

vis_dat(typical_data)

```

We can also look into using even wider data, looking at `typical_larger_data`

```{r vis-dat-typical-larger-data}

vis_dat(typical_data_large)

```

## Using `vis_miss()`

We can explore the missing data further using `vis_miss()`.

```{r vis-miss-aq}

vis_miss(airquality)

```

The percentages of missing/complete in `vis_miss` are accurate to 1 decimal place.

You can cluster the missingness by setting `cluster = TRUE`.

```{r vis-miss-aq-cluster}

vis_miss(airquality, 
         cluster = TRUE)

```

The columns can also just be arranged by columns with most missingness, by setting `sort_miss = TRUE`.

```{r vis-miss-aq-sort-miss}

vis_miss(airquality,
         sort_miss = TRUE)

```

`vis_miss` indicates when there is a very small amount of missing data at <0.1% missingness.

```{r vis-miss-test}

test_miss_df <- data.frame(x1 = 1:10000,
                           x2 = rep("A", 10000),
                           x3 = c(rep(1L, 9999), NA))

vis_miss(test_miss_df)

```

`vis_miss` will also indicate when there is no missing data at all.

```{r vis-miss-mtcars}

vis_miss(mtcars)

```

# Thank yous

Thank you to Ivan Hanigan who [first commented](http://www.njtierney.com/r/missing%20data/rbloggers/2015/12/01/ggplot-missing-data/#comment-2388263747) this suggestion after I made a blog post about an initial prototype `ggplot_missing`, and Jenny Bryan, whose [tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me thinking about `vis_dat`, and for her code contributions that removed a lot of errors.

Thank you to Hadley Wickham for suggesting the use of the internals of `readr` to make `vis_guess` work.

Thank you to Miles McBain for his suggestions on how to improve `vis_guess`. This resulted in making it at least 2-3 times faster.

Thanks to Carson Sievert for writing the code that combined `plotly` with `visdat`, and for Noam Ross for suggesting this in the first place.

Finally thank you to [rOpenSci](https://github.com/ropensci) and it's amazing [onboarding process](https://github.com/ropensci/onboarding), this process has made visdat a much better package, thanks to the editor Noam Ross (@noamross), and the reviewers Sean Hughes (@seaaan) and Mara Averick (@batpigandme).

[![ropensci_footer](https://ropensci.org/public_images/ropensci_footer.png)](https://ropensci.org)