### 2023-02-01 README.Rmd
check URLs

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

```

# visdat <img src="man/figures/visdat-logo.png" align="right" />

<!-- badges: start -->
[![rOpenSci Badge](https://badges.ropensci.org/87_status.svg)](https://github.com/ropensci/software-review/issues/87)[![JOSS status](https://joss.theoj.org/papers/10.21105/joss.00355/status.svg)](https://joss.theoj.org/papers/10.21105/joss.00355)[![DOI](https://zenodo.org/badge/50553382.svg)](https://zenodo.org/badge/latestdoi/50553382)[![R-CMD-check](https://github.com/ropensci/visdat/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ropensci/visdat/actions/workflows/R-CMD-check.yaml)[![Codecov test coverage](https://codecov.io/gh/ropensci/visdat/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ropensci/visdat?branch=master)[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/visdat)](https://cran.r-project.org/package=visdat)[![CRAN Logs](http://cranlogs.r-pkg.org/badges/visdat)](https://CRAN.R-project.org/package=visdat)[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/)
<!-- badges: end -->

# How to install

visdat is available on CRAN

```{r install-cran, eval = FALSE}

install.packages("visdat")

```

If you would like to use the development version, install from github with:

```{r installation, eval = FALSE}

# install.packages("devtools")
devtools::install_github("ropensci/visdat")
```

# What does visdat do?

Initially inspired by
[`csv-fingerprint`](https://github.com/setosa/csv-fingerprint), `vis_dat` helps
you visualise a dataframe and "get a look at the data" by displaying the
variable classes in a dataframe as a plot with `vis_dat`, and getting a brief
look into missing data patterns using `vis_miss`.

`visdat` has 6 functions:

- `vis_dat()` visualises a dataframe showing you what the classes of the columns
are, and also displaying the missing data.

- `vis_miss()` visualises just the missing data, and allows for missingness to
be clustered and columns rearranged. `vis_miss()` is similar to
`missing.pattern.plot` from the
[`mi`](https://CRAN.R-project.org/package=mi) package.
Unfortunately `missing.pattern.plot` is no longer in the `mi` package (as of
14/02/2016).

- `vis_compare()` visualise differences between two dataframes of the same 
  dimensions

- `vis_expect()` visualise where certain conditions hold true in your data

- `vis_cor()` visualise the correlation of variables in a nice heatmap

- `vis_guess()` visualise the individual class of each value in your data
- `vis_value()` visualise the value class of each cell in your data
- `vis_binary()` visualise the occurrence of binary values in your data

You can read more about visdat in the vignette, ["using visdat"]https://docs.ropensci.org/visdat/articles/using_visdat.html).

## Code of Conduct

Please note that the visdat project is released with a [Contributor Code of Conduct](https://github.com/ropensci/visdat/blob/master/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

# Examples

## Using `vis_dat()`

Let's see what's inside the `airquality` dataset from base R, which contains
information about daily air quality measurements in New York from May to
September 1973. More information about the dataset can be found with
`?airquality`.

```{r vis-dat-aq}

library(visdat)

vis_dat(airquality)

```

The plot above tells us that R reads this dataset as having numeric and integer
values, with some missing data in `Ozone` and `Solar.R`. The classes are
represented on the legend, and missing data represented by grey. The
column/variable names are listed on the x axis.

The `vis_dat()` function also has a `facet` argument, so you can create small multiples of a similar plot for a level of a variable, e.g., Month:

```{r vis-dat-month}
vis_dat(airquality, facet = Month)
```

These currently also exist for `vis_miss()`, and the `vis_cor()` functions.

## Using `vis_miss()`

We can explore the missing data further using `vis_miss()`:

```{r vis-miss-aq}

vis_miss(airquality)

```

Percentages of missing/complete in `vis_miss` are accurate to the integer (whole number). To get more accurate and thorough exploratory summaries of missingness, I would recommend the [`naniar`](https://github.com/njtierney/naniar) R package

You can cluster the missingness by setting `cluster = TRUE`:

```{r vis-miss-aq-cluster}

vis_miss(airquality, 
         cluster = TRUE)

```

Columns can also be arranged by columns with most missingness, by setting
`sort_miss = TRUE`:

```{r vis-miss-aq-sort-miss}

vis_miss(airquality,
         sort_miss = TRUE)

```

`vis_miss` indicates when there is a very small amount of missing data at <0.1%
missingness:

```{r vis-miss-test}

test_miss_df <- data.frame(x1 = 1:10000,
                           x2 = rep("A", 10000),
                           x3 = c(rep(1L, 9999), NA))

vis_miss(test_miss_df)

```

`vis_miss` will also indicate when there is no missing data at all:

```{r vis-miss-mtcars}

vis_miss(mtcars)

```

To further explore the missingness structure in a dataset, I recommend the
[`naniar`](https://github.com/njtierney/naniar) package, which provides more
general tools for graphical and numerical exploration of missing values.

## Using `vis_compare()`

Sometimes you want to see what has changed in your data. `vis_compare()` displays the differences in two dataframes of the same size. Let's look at an example.

Let's make some changes to the `chickwts`, and compare this new dataset:

```{r vis-compare-iris}
set.seed(2019-04-03-1105)
chickwts_diff <- chickwts
chickwts_diff[sample(1:nrow(chickwts), 30),sample(1:ncol(chickwts), 2)] <- NA

vis_compare(chickwts_diff, chickwts)

```

Here the differences are marked in blue.

If you try and compare differences when the dimensions are different, you get 
an ugly error:

```{r vis-compare-error, eval = FALSE}

chickwts_diff_2 <- chickwts
chickwts_diff_2$new_col <- chickwts_diff_2$weight*2

vis_compare(chickwts, chickwts_diff_2)
# Error in vis_compare(chickwts, chickwts_diff_2) : 
#   Dimensions of df1 and df2 are not the same. vis_compare requires dataframes of identical dimensions.
```

## Using `vis_expect()`

`vis_expect` visualises certain conditions or values in your data. For example,
If you are not sure whether to expect values greater than 25 in your data
(airquality), you could write: `vis_expect(airquality, ~.x>=25)`, and you can
see if there are times where the  values in your data are greater than or equal
to 25:

```{r vis-expect}

vis_expect(airquality, ~.x >= 25)

```

This shows the proportion of times that there are values greater than 25, as
well as the missings.

## Using `vis_cor()`

To make it easy to plot correlations of your data, use `vis_cor`:

```{r vis-cor}

vis_cor(airquality)

```

## Using `vis_value`

`vis_value()` visualises the values of your data on a 0 to 1 scale.

```{r vis-value}
vis_value(airquality)
```

It only works on numeric data, so you might get strange results if you are using   factors:

```{r iris-error, eval = FALSE}
library(ggplot2)
vis_value(iris)
```

```
data input can only contain numeric values, please subset the data to the numeric values you would like. dplyr::select_if(data, is.numeric) can be helpful here!
```

So you might need to subset the data beforehand like so:

```{r iris-error-fix}
library(dplyr)

iris %>%
  select_if(is.numeric) %>%
  vis_value()
```

## Using `vis_binary()`

`vis_binary()` visualises binary values. See below for use with example data, `dat_bin`

```{r vis-bin}
vis_binary(dat_bin)
```

If you don't have only binary values a warning will be shown.

```{r vis-bin-airq, eval = FALSE}
vis_binary(airquality)
```

```
Error in test_if_all_binary(data) : 
  data input can only contain binary values - this means either 0 or 1, or NA. Please subset the data to be binary values, or see ?vis_value.
```

## Using `vis_guess()`

`vis_guess()` takes a guess at what each cell is. It's best illustrated using
some messy data, which we'll make here:

```{r create-messy-vec}

messy_vector <- c(TRUE,
                  T,
                  "TRUE",
                  "T",
                  "01/01/01",
                  "01/01/2001",
                  NA,
                  NaN,
                  "NA",
                  "Na",
                  "na",
                  "10",
                  10,
                  "10.1",
                  10.1,
                  "abc",
                  "$%TG")

set.seed(2019-04-03-1106)
messy_df <- data.frame(var1 = messy_vector,
                       var2 = sample(messy_vector),
                       var3 = sample(messy_vector))

```

```{r vis-guess-messy-df, fig.show='hold', out.width='50%'}

vis_guess(messy_df)
vis_dat(messy_df)

```

So here we see that there are many different kinds of data in your dataframe. As
an analyst this might be a depressing finding. We can see this comparison above.

# Thank yous

Thank you to Ivan Hanigan who [first
commented](https://www.njtierney.com/post/2015/11/12/ggplot-missing-data/)
this suggestion after I made a blog post about an initial prototype
`ggplot_missing`, and Jenny Bryan, whose
[tweet](https://twitter.com/JennyBryan/status/679011378414268416) got me
thinking about `vis_dat`, and for her code contributions that removed a lot of
errors.

Thank you to Hadley Wickham for suggesting the use of the internals of `readr`
to make `vis_guess` work. Thank you to Miles McBain for his suggestions on how
to improve `vis_guess`. This resulted in making it at least 2-3 times faster.
Thanks to Carson Sievert for writing the code that combined `plotly` with
`visdat`, and for Noam Ross for suggesting this in the first place. Thank you
also to Earo Wang and Stuart Lee for their help in getting capturing expressions
in `vis_expect`.

Finally thank you to [rOpenSci](https://github.com/ropensci) and it's amazing
[onboarding process](https://github.com/ropensci/software-review), this process has
made visdat a much better package, thanks to the editor Noam Ross (@noamross),
and the reviewers Sean Hughes (@seaaan) and Mara Averick (@batpigandme).

[![ropensci_footer](https://ropensci.org/public_images/ropensci_footer.png)](https://ropensci.org)